### 2020-06-17 report/sections/appendix/eval.tex
Report: new conclusion
Report: feedback

@@

\section{Implementation for \tprotocol{Two Buyer}}
\label{section:evalcodetwobuyer}

\section{Implementation for \tprotocol{ATM}}
\label{section:evalcodeatm}

>>>>

\begin{lstlisting}[language=javascript,tabsize=2]
/**
 * User implementation of game board state.
 * Export game board instance.
 */

import { Coordinate as Point } from './Game/GameTypes';

enum Cell { Empty, P1, P2 };
export enum MoveResult { Win, Draw, Continue };

class Board {

	private _board: Array<Array<Cell>>;
	private _emptyCellCount: number;

	constructor() {
		this._board = [[Cell.Empty, Cell.Empty, Cell.Empty],
		[Cell.Empty, Cell.Empty, Cell.Empty],
		[Cell.Empty, Cell.Empty, Cell.Empty]];
		this._emptyCellCount = 9;
	}

	// Factory for generating a function that, given a player role,
	// places the marker and returns the game result with respect to 
	// that player.
	private _makeMove = (marker: Cell) => ({ x: row, y: col }: Point) 
		=> {
		return new Promise<MoveResult>((resolve, _) => {
			setTimeout(() => {
				// Update board state
				this._board[row][col] = marker;
				this._emptyCellCount--;

				// Check for winning move
				if (this._board[row].every(cell => cell === marker))
					resolve(MoveResult.Win);

				/* Winning column */
				if (this._board.every(row => row[col] === marker))
					resolve(MoveResult.Win);

				/* Placed middle marker - check corners */
				if (row === 1 && col === 1) 
					if ((this._board[0][0] === marker 
						&& this._board[2][2] === marker) ||
						(this._board[0][2] === marker 
						&& this._board[2][0] === marker))
						resolve(MoveResult.Win);

				/* Placed corner marker - check diagonals */
				if (row !== 1 && col !== 1) 
					if (this._board[1][1] === marker 
						&& this._board[2 - row][2 - col] === marker)
						resolve(MoveResult.Win);

				resolve(this._emptyCellCount === 0 ? 
					MoveResult.Draw : MoveResult.Continue);
			}, 2000);
		})

	}

	p1(move: Point) { return this._makeMove(Cell.P1)(move); }
	p2(move: Point) { return this._makeMove(Cell.P2)(move); }

	clear() {
		this._board = [[Cell.Empty, Cell.Empty, Cell.Empty],
		[Cell.Empty, Cell.Empty, Cell.Empty],
		[Cell.Empty, Cell.Empty, Cell.Empty]];
		this._emptyCellCount = 9;
	}

}

// Initialise state
export const board = new Board();
\end{lstlisting}

@@

%\section{Implementation for \tprotocol{Two Buyer}}
%\label{section:evalcodetwobuyer}

\section{Package Dependencies for Benchmarks}
\label{section:evalbenchmarkpackagejson}

\begin{lstlisting}[language=javascript,tabsize=2,
title=Packages for Ping Pong Client]
{
  "dependencies": {
    "@testing-library/jest-dom": "^4.2.4",
    "@testing-library/react": "^9.3.2",
    "@testing-library/user-event": "^7.1.2",
    "@types/jest": "^24.0.0",
    "@types/node": "^12.0.0",
    "@types/react": "^16.9.0",
    "@types/react-dom": "^16.9.0",
    "react": "^16.13.1",
    "react-dom": "^16.13.1",
    "react-scripts": "3.4.1",
    "typescript": "~3.7.2"
  },
  "devDependencies": {
    "@babel/preset-typescript": "^7.10.1",
    "customize-cra": "^1.0.0",
    "react-app-rewired": "^2.1.6"
  }
}
\end{lstlisting}

\begin{lstlisting}[language=javascript,tabsize=2,
title=Packages for Ping Pong Server]
{
  "dependencies": {
    "express": "^4.17.1",
    "ws": "^7.3.0",
    "zombie": "^6.1.4"
  },
  "devDependencies": {
    "@types/express": "^4.17.6",
    "@types/ws": "^7.2.4",
    "ts-node": "^8.10.2",
    "typescript": "^3.9.3"
  }
}
\end{lstlisting}

### 2020-06-17 report/sections/evalconcl/eval.tex
Report: feedback

In this chapter, we
evaluate the expressiveness
and performance of our API generation
toolchain.

@@

To evaluate the expressiveness
of our work,
we use two case studies
of protocols found in webservices
to demonstrate strengths and weaknesses
of our work:

@@

\item
\textbf{\tprotocol{Noughts and Crosses} Game:}

In \cref{section:evalgame},
we implement a classic multiplayer game which involves a game server
and two players interacting with the game via the browser,
to show that our work is compatible with multiparty sessions,
which isn't the case in \cite{MVU2020,Exceptional}.
We show how our generated APIs are compatible with the
state management solution used by the game, and give examples of 
how the generated APIs empower the developer to intuitively 
implement the game interface.

\item
\textbf{\tprotocol{Two Buyers} Protocol, as introduced in \cref{section:twobuyer}:}

In \cref{section:evaltwobuyer},
we walk through an implementation of the \tprotocol{Two Buyers}
protocol. This \textit{cannot} be implemented in
\cite{PureScript2019,MVU2020}, which illustrates the novelty
of our \newtheory theory on routed multiparty session types.
We emphasise how the routing mechanism is transparent to the
developer implementing the protocol, which demonstrates
the practicality of our extensions to \codegen to support
\newtheory.

@@

To evaluate the performance of our work,
we run micro-benchmarks
on a variant of the \tprotocol{Ping Pong}
protocol parameterised by the number
of round trips
to analyse the overhead of our implementation,
compared with implementations written
without using our generated APIs,
as the number of round trips increase.
We describe the experiment methodology
and comment on our findings in \cref{section:benchmarks}.

>>>>

In this chapter, we
evaluate the expressiveness
and performance of our generated APIs.

@@

\subsubsection{Expressiveness}
To evaluate the expressiveness
of our work,
we use two case studies
of protocols found in webservices
to demonstrate strengths and weaknesses
of our work:

@@

\item
\textbf{\tprotocol{Noughts and Crosses} Game, as introduced
in \cite{PLACES2020}:}

In \cref{section:evalgame},
we implement a classic multiplayer game which involves a game server
and two players interacting with the game via the browser,
to show that our work is compatible with multiparty sessions,
which is not the case in \cite{MVU2020,Exceptional}.
We show how our generated APIs are compatible with the
state management solution used by the game, and give examples of 
how the generated APIs empower the developer to intuitively 
implement the game interface.

\item
\textbf{\tprotocol{Two Buyers} Protocol, as introduced in 
\cref{section:twobuyer} and \cite{JACM2016,MPAST}:}

In \cref{section:evaltwobuyer},
we walk through an implementation of the \tprotocol{Two Buyers}
protocol. This \textit{cannot} be implemented in
\cite{PureScript2019,MVU2020,Exceptional}, 
which illustrates the novelty
of our \newtheory theory on routed multiparty session types.
We emphasise how the routing mechanism is transparent to the
developer implementing the protocol, which demonstrates
the practicality of our extensions to \codegen to support
\newtheory.

@@

\subsubsection{Performance}
To evaluate the performance of our work,
we run micro-benchmarks
on a variant of the \tprotocol{Ping Pong}
protocol parameterised by the number
of round trips
to analyse the overhead of our implementation,
compared with implementations written
without using our generated APIs,
as the number of round trips increase.
We describe the experiment methodology
and comment on our findings in \cref{section:benchmarks}.

### 2020-06-17 report/sections/evalconcl/eval/benchmarks.tex
Report: complete chapter 10.1
Report: feedback

\section{Performance Benchmarks}

@@

\item We use the production build generated by 
\texttt{create-react-app} \cite{cra} for all experiments, which performs the
transpilation into JavaScript. We serve the production build using the
\texttt{serve} package \cite{npmserve} available on \texttt{npm}.

@@

\item All \trole{Svr}s run the benchmarks without a real web browser,
using headless browsing functionality from 
the Zombie.js \cite{zombie} package.

@@

We run the experiments under a network of latency 0.165ms
(64 bytes ping), and repeat each experiment 20 times.
Execution time measurements  are taken using a machine 
equipped with Intel i7-4850HQ CPU (2.3 GHz, 4 cores, 8 threads), 
16 GB RAM, macOS operating system version 10.15.4, 
Node.js runtime version 12.12.0, and
TypeScript compiler version 3.7.4.
We standardise all packages used in the front- and back-end
implementations across experiments. Details can be found in their
corresponding \texttt{package.json} manifests (\cref{appendix:eval}).

>>>>

\section{Performance Benchmarks}
\label{section:benchmarks}

@@

\item We use the production build generated by 
\textit{create-react-app} \cite{cra} for all experiments, which performs the
transpilation into JavaScript. We serve the production build using the
\textit{serve} package \cite{npmserve} available on \textit{npm}.

@@

\item All \trole{Svr}s run the benchmarks without a real web browser,
using headless browsing functionality from 
the \textit{Zombie.js} \cite{zombie} package.

@@

We run the experiments under a network of latency 0.165ms
(64 bytes ping), and repeat each experiment 20 times.
Execution time measurements  are taken using a machine 
equipped with Intel i7-4850HQ CPU (2.3 GHz, 4 cores, 8 threads), 
16 GB RAM, macOS operating system version 10.15.4, 
Node.js runtime version 12.12.0, and
TypeScript compiler version 3.7.4.
We standardise all packages used in the front- and back-end
implementations across experiments. Details can be found in their
corresponding \texttt{package.json} manifests 
(\cref{section:evalbenchmarkpackagejson}).

### 2020-06-17 report/sections/evalconcl/eval/game.tex
Report: feedback

We implement the classic turn-based board game of \textit{Noughts and Crosses}
between two players,
as introduced in \cref{section:intro}.
We formalise the game interactions using a Scribble protocol
presented in \cref{lst:game}.
The \tprotocol{Game} protocol describes one turn:
\trole{P1} makes a move by sending the coordinates of a
vacant cell on the game board to \trole{Svr},
then \trole{Svr} reports the outcome of that move
to both players. If another round is required
to determine the game result, the \tprotocol{Game}
protocol is recursively invoked (\cref{line:gamerec}) with
roles \trole{P1} and \trole{P2} swapped.

@@

\subsection{Game Server}
We set up the WebSocket server as an Express.js \cite{Express} application
on top of the Node.js runtime.
We define our own game logic in a \texttt{Board} class
to keep track of the game state and expose methods 
to query the result -- the implementation for \texttt{Board} is included
in \cref{section:evalcodegame}.
This custom logic is integrated into the 
\texttt{handleP1Move} and \texttt{handleP2Move}
handlers implemented by the developer, defined to
handle the moves made by \trole{P1} and \trole{P2} respectively.
We illustrate this in \cref{lst:gamesvr}.

\begin{figure}[!h]
\begin{lstlisting}[language=javascript,tabsize=2]
const handleP1Move = new Implementation.S13({
	Pos: async (move: Point) => {
		// `board` manages game state;
		// `board.P1` registers the move and returns the game result
		const result = await board.P1(move);
		switch (result) {
			case MoveResult.Win: {
				return new Implementation.S15([
					// Send losing result to P2
					[Labels.S15.Lose, [move], new Implementation.S16(
						// Send winning result to P1
						[Labels.S16.Win, [move], new Implementation.Terminal()]
					)]
				]);
			}
			case MoveResult.Draw: { ... }
			case MoveResult.Continue: {
				return new Implementation.S15(
					// Notify both players and proceed
					// with next round using `handleP2Move`
					[Labels.S15.Update, [move], new Implementation.S18(
						[Labels.S18.Update, [move], handleP2Move]
					)]
				);	
			}		
		}
	}
});

@@

new Svr(wss, handleP1Move, (role, reason) => {
	console.log(`${role} cancelled because of ${reason}`);
});
\end{lstlisting}
\captionof{lstlisting}{Implementing \tprotocol{Noughts and Crosses} Game Server using
\codegen}
\label{lst:gamesvr}
\end{figure}

@@

The main implementation detail for players
is to make a move. 
Intuitively, the developer implements a grid and binds a handler
to the \lstonelinejs{'onClick'} event of each vacant cell to send that cell's coordinate 
in a \tmsg{Pos(Point)} message to the game server.
A common source of bugs would be not preventing the user from selecting
a second cell when waiting for the game server's response,
which violates the game rules (and the global protocol).

@@

>>>>

We implement the classic turn-based board game of 
\textit{Noughts and Crosses}
between two players,
as introduced in \cite{PLACES2020}.
Both players, identified by either \textit{noughts (O's)}
or \textit{crosses (X's)} respectively, take turns
to place a mark on an unoccupied cell of a 3-by-3 grid
until one player wins (when their markers form one straight
line on the board) or a stalemate is reached (when all cells
are occupied and no one wins).

We formalise the game interactions using a Scribble protocol
presented in \cref{lst:game}.
The \tprotocol{Game} protocol describes one turn:
\trole{P1} makes a move by sending the coordinates of a
vacant cell on the game board to \trole{Svr},
then \trole{Svr} reports the outcome of that move
to both players. If another round is required
to determine the game result, the \tprotocol{Game}
protocol is recursively invoked (\cref{line:gamerec}) with
roles \trole{P1} and \trole{P2} swapped.

@@

\subsection{Game Server}
We set up the WebSocket server as an \textit{Express.js}\footnote{
Express is a commonly used library for writing lightweight
web servers in JavaScript.
} \cite{Express} application
on top of the Node.js runtime.
We define our own game logic in a \texttt{Board} class
to keep track of the game state and expose methods 
to query the result -- the implementation for \texttt{Board} is included
in \cref{section:evalcodegame}.
This custom logic is integrated into the 
\texttt{handleP1Move} and \texttt{handleP2Move}
handlers implemented by the developer, defined to
handle the moves made by \trole{P1} and \trole{P2} respectively.
We illustrate this in \cref{lst:gamesvr}.

\begin{figure}[!h]
\begin{lstlisting}[language=javascript,tabsize=2]
const handleP1Move = new Implementation.S13({
	Pos: (*@\hl{async}@*) (move: Point) => {
		// `board` manages game state;
		// `board.P1` registers the move and returns the game result
		const result = (*@\hl{await}@*) board.P1(move);
		switch (result) {
			case MoveResult.Win: {
				return new Implementation.S15([
					// Send losing result to P2
					[Labels.S15.Lose, [move], new Implementation.S16(
						// Send winning result to P1
						[Labels.S16.Win, [move], new Implementation.Terminal()]
					)]
				]);
			}
			case MoveResult.Draw: { ... }
			case MoveResult.Continue: {
				return new Implementation.S15(
					// Notify both players and proceed
					// with next round using `handleP2Move`
					[Labels.S15.Update, [move], new Implementation.S18(
						[Labels.S18.Update, [move], handleP2Move]
					)]
				);	
			}		
		}
	}
});

@@

const cancellation = (role: Roles.All, reason: string) => {
	console.log(`${role} cancelled because of ${reason}`);
}

new Svr(
	wss,						// WebSocket server
	handleP1Move,		// Game logic
	cancellation		// Error handler
);
\end{lstlisting}
\captionof{lstlisting}{Implementing \tprotocol{Noughts and Crosses} Game Server using
\codegen}
\label{lst:gamesvr}
\end{figure}

@@

The main implementation detail for players
is to make moves. 
Intuitively, the developer implements a grid and binds a handler
to the \lstonelinejs{'onClick'} event of each vacant cell to send that cell's coordinate 
in a \tmsg{Pos(Point)} message to the game server.
A common source of bugs would be not preventing the user from selecting
a second cell when waiting for the game server's response,
which violates the game rules (and the global protocol).

@@

Code is available at 
\url{https://github.com/ansonmiu0214/SessionTS-Examples/NoughtsAndCrosses}.

### 2020-06-17 report/sections/evalconcl/eval/twobuyer.tex
Report: feedback

We implement the \tprotocol{Two Buyer} protocol
introduced in \cref{section:twobuyer}. 
This protocol cannot be implemented using
existing proposals \cite{PureScript2019,MVU2020} for integrating
session types into web development.
\codegen overcomes the limitations of existing work
through implementing novel theory of routed multiparty
session types formalised in \cref{chap:theory}.

@@

\begin{figure}[!h]
\begin{lstlisting}[language=javascript]
// two buyer seller
\end{lstlisting}
\captionof{lstlisting}{\dots}
\label{lst:evaltwobuyerseller}
\end{figure}

\begin{figure}[!h]
\begin{lstlisting}[language=javascript]
// two buyer A p2p
\end{lstlisting}
\captionof{lstlisting}{\dots}
\label{lst:evaltwobuyerA}
\end{figure}

The \newtheory implementation does not affect
the compatibility of our generated APIs with external libraries.
The \trole{S}eller endpoint is set up as an Express.js
application, and both buyers still use the React Context API
for application state management. We show an example of
the latter in the developer's implementation for
buyer \trole{A} receiving and keeping track of the server's quote
using a React Context in \cref{lst:evaltwobuyercontext}.

\begin{figure}[!h]
\begin{lstlisting}[language=javascript]
// two buyer A context
\end{lstlisting}
\captionof{lstlisting}{\dots}
\label{lst:evaltwobuyercontext}
\end{figure}

We include the implementations of the remaining EFSM states
for buyer \trole{A} in \cref{section:evalcodetwobuyer}
for completeness;
interested readers can find the full implementations for the other
endpoints alongside the generated code on
GitHub\footnote{
\url{https://github.com/ansonmiu0214/SessionTS-Examples/TwoBuyer}
}.

>>>>

We implement the \tprotocol{Two Buyer} protocol
introduced in \cref{section:twobuyer}. 
This protocol cannot be implemented using
existing proposals \cite{PureScript2019,MVU2020} for integrating
session types into web development.
\codegen overcomes the limitations of existing work
through implementing novel theory of routed multiparty
session types formalised in \cref{chap:theory};
the implementation of \newtheory in \codegen is 
explained in \cref{chap:impl}.

@@

\begin{figure}[!h]
\begin{lstlisting}[language=javascript]
new Implementation.Initial({
	[Labels.S32.title]: (title) => {
		return new Implementation.S34([
			Labels.S34.quote, [100], new Implementation.S35([
				Labels.S35.quote, [100], new Implementation.S36({
					[Labels.S36.buy]: () => {
						return new Implementation.Terminal();
					},
					[Labels.S36.cancel]: () => {
						return new Implementation.Terminal();
					}
				})
			])
		])
	}
})
\end{lstlisting}
\captionof{lstlisting}{Two Buyer Seller Implementation}
\label{lst:evaltwobuyerseller}
\end{figure}

\begin{figure}[!h]
\begin{lstlisting}[language=javascript]
type State = { split: number };

class ProposeSplit extends S11<State> {
	state = { split: 0 };

	render() {
		const SendSplit = this.split('onClick', ev => {
			ev.preventDefault();
			return [this.state.split];		
		});

		return <div>
			<input
				type='number'
				value={this.state.split}
				onChange={ev => this.setState({
					split: Number(ev.target.value)
				})}
				placeholder='Enter split'
			/>
			<SendSplit><button>Propose</button></SendSplit>	
		</div>;	
	}
}
\end{lstlisting}
\captionof{lstlisting}{Developer Implementation of Peer-to-Peer Interaction
in Two Buyer}
\label{lst:evaltwobuyerA}
\end{figure}

The \newtheory implementation does not affect
the compatibility of our generated APIs with external libraries.
The \trole{S}eller endpoint is set up as an Express.js
application, and both buyers use the \textit{React Context API}\footnote{
React Contexts allow components
to pass data (such as application state)
through the component tree without
having to propagate via props at every level.
}
for application state management. 
%We show an example of
%the latter in the developer's implementation for
%buyer \trole{A} receiving and keeping track of the server's quote
%using a React Context in \cref{lst:evaltwobuyercontext}.
%
%\begin{figure}[!h]
%\begin{lstlisting}[language=javascript]
%// two buyer A context
%\end{lstlisting}
%\captionof{lstlisting}{\dots}
%\label{lst:evaltwobuyercontext}
%\end{figure}

%We include the implementations of the remaining EFSM states
%for buyer \trole{A} in \cref{section:evalcodetwobuyer}
%for completeness;
Interested readers can find the full implementations for the other
endpoints alongside the generated code on
GitHub\footnote{
\url{https://github.com/ansonmiu0214/SessionTS-Examples/TwoBuyer}
}.

### 2020-06-17 report/sections/general/implementation.tex
Report: feedback

\chapter{Implementation}
\label{chap:impl}

@@

We can now adapt the \texttt{onmessage}
WebSocket event listener defined at the server endpoint
to route messages not intended for the server role,
as shown in \cref{lst:noderoutereceive}.
By construction from our API generation strategy,
the incoming message must match the interface defined
above, so it is a safe deserialisation step.
This preliminary deserialisation allows the server
to identify \textit{routed messages}, and proceed to
send it to the intended recipient.

@@

Hence, we need to implement
the message receiving framework outlined in \cref{subsection:noderuntimereceive}
instead, which generalises the pair of message/handler queues to two mappings
to provide a pair of message/handler queues for each endpoint.
The runtime instantiates the two mappings of queues,
and receive handler registration now targets the pair of queues
specified by the role from which to receive from.
We illustrate this in \cref{lst:reactrouterreceive}

>>>>

\chapter{Implementing \newtheory in \codegen}
\label{chap:impl}

@@

We can now adapt the \texttt{onmessage}
WebSocket event listener defined at the server endpoint
to route messages not intended for the server role,
as shown in \cref{lst:noderoutereceive}.
By construction of our API generation strategy,
the incoming message must match the interface defined
above, so it is a safe deserialisation step.
This preliminary deserialisation allows the server
to identify \textit{routed messages}, and proceed to
send it to the intended recipient.

@@

Hence, we need to implement
the message receiving framework outlined in 
\cref{subsection:noderuntimereceive}
instead, which generalises the pair of message/handler queues to two mappings
to provide a pair of message/handler queues for each endpoint.
The runtime instantiates the two mappings of queues,
and receive handler registration now targets the pair of queues
specified by the role to receive from.
We illustrate this in \cref{lst:reactrouterreceive}

### 2020-06-17 report/sections/general/p2p.tex
Report: feedback

In this chapter,
we demonstrate how to support peer-to-peer interactions
between web browser endpoints over WebSocket transport,
thus relaxing the server-centric topology assumption from 
\cref{part:server}.

>>>>

In this chapter,
we demonstrate how to support peer-to-peer interactions
through routing communications
between web browser endpoints over WebSocket transport,
thus relaxing the server-centric topology assumption from 
\cref{part:server}.

### 2020-06-17 report/sections/general/p2p/challenges.tex
Report: feedback

We need to formalise our concept of routing by extending session
type theory, such that, given a communication protocol implementing 
an arbitrary topology, we can encode it into our extended theory and
prove that safety properties (e.g. well-formedness, deadlock freedom)
are preserved, and that communication traces are preserved.
This must be done \textit{very carefully}, as
naive definitions risk over-serialising the original communication,
thereby not preserving all communication traces.

>>>>

We need to formalise our concept of routing by extending session
type theory, such that, given a communication protocol implementing 
an arbitrary topology, we can encode it into our extended theory and
prove that safety properties (e.g. well-formedness, deadlock freedom)
are preserved, and that communication traces are preserved.
This must be done \textit{very carefully}, as
naive definitions of routing risk 
over-serialising the original communication,
thereby not preserving all communication traces.

### 2020-06-17 report/sections/general/p2p/twobuyer.tex
Report: feedback

We present the \tprotocol{Two Buyer} protocol in \cref{lst:twobuyer},
which describes a protocol between two buyers \trole{A}, \trole{B}
and a \trole{S}eller. We assume that the \trole{S}eller runs on
the server, and the buyers run on the browser.

>>>>

We present the \tprotocol{Two Buyer} protocol in \cref{lst:twobuyer}
-- the canonical example for peer-to-peer interactions in
literature \cite{JACM2016,MPAST} --
which describes a protocol between two buyers \trole{A}, \trole{B}
and a \trole{S}eller.
We assume that the \trole{S}eller runs on
the server, and the buyers run on the browser.

### 2020-06-17 report/sections/server/extensions/async.tex
Report: feedback

Asynchronous APIs are commonplace in modern web programming,
ranging from database operations to third-party API calls.
We motivate the need for supporting asynchronous operations
through an authentication protocol in
\cref{subsection:asyncmotivation},
demonstrate how we support this through extending
the generated APIs and runtime in
\cref{subsection:asyncapi,subsection:asyncruntime}
respectively, 
and highlight the limitations of our approach in
\cref{subsection:asynclimit}.

@@

The developer's implementation on 
\cref{line:2fachoice1,line:2fachoice2} is likely to query
a database for credential verification.
Database operations are implemented using \textit{asynchronous},
\textit{non-blocking} APIs; it is
recommended that file operations do not block the main execution,
so a JavaScript database API may require a \textit{callback}
to propagate the return value when it is available.
A server endpoint implementation of \cref{lst:2FA}
may resemble the following:

@@

Callbacks are commonplace in dealing with concurrency in JavaScript
\cite{CallbackHell}; we will go over other widely-used
concurrency primitives shortly.
Whatever the case may be, asynchronous developer implementations
using our generated APIs will not type-check as the function
returns immediately (hence, has return type \lstonelinejs{void}),
even though the callback function has the expected return type.
The developer will encounter the following:

@@

This is also possible for browser-side implementations:
when the user invokes a UI event, the browser-side logic
may first consult a third-party API for additional information
(which is not part of the communication protocol)
before performing a send action; the \textit{Fetch} API \cite{Fetch}
for making API calls is asynchronous.

@@

This means we should support
the concurrency primitives built into TypeScript:

\subparagraph{Callbacks}
Callback-based APIs are higher-order functions -- they take functions
as parameters and invoke them when the asynchronous operation is complete.
The \texttt{setTimeout(fn, delay)} function available on the browser
waits for \texttt{delay} milliseconds before invoking \texttt{fn()} --
or more precisely, before queuing the execution of \texttt{fn()}.
As a result, the code listing below prints \lstonelinejs{'Second!'}
before \lstonelinejs{'First!'}, even though \texttt{setTimeout}
was called with 0ms delay.

@@

\subparagraph{Promises}
A Promise formalises the ``completion'' callback
and its error handling construct. The developer passes a ``success''
handler and ``error'' handler, conventionally named
``then'' and ``catch'' respectively, to construct a Promise.
This is analogue to the \texttt{Future} interface in Java \cite{JavaFuture}.
The Promise will invoke the correct handler on completion.
The benefits of Promises in our context is that TypeScript allows
us to specify the type of payload that the Promise resolves to.

@@

This is rather straightforward.
We construct a generic union type, \texttt{MaybePromise<T>},
to specify that the type parameter might be wrapped in a Promise,
and leverage type inference in conditional types to
define a generic type, \texttt{FromPromise<T>}, to extract
the wrapped type.

@@

Now we change the type signatures of our generated APIs
to use \texttt{MaybePromise<T>}.
For example, if the handler of a receive state generated by
\fancyname{NodeMPST} must return \texttt{Implementation.S42},
it is now permitted to return a \texttt{Promise} that resolves
to a value of type texttt{Implementation.S42}.
The same logic applies to other types of EFSM states
in both \fancyname{NodeMPST} and \fancyname{ReactMPST}.

Hence, the asynchronous \trole{Svr} implementation
for the 2FA protocol shown in \cref{lst:new2fa}
now type-checks. Note that, because 
\lstonelinejs{async} functions return \texttt{Promise}s,
we also support this newer piece of syntactic sugar.

@@

Generally, minimal changes are required for the runtime.
The key observation is that the concrete types for the any
handlers are the same -- they just might be wrapped in a Promise
construct.
To handle potential Promises, we wrap existing logic
as a delayed computation, colloquially known as a \textit{thunk}
\cite{Thunk}.
We show an example in \cref{fig:thunk} of how we
adapt the \texttt{performSend()} function generated by
\fancyname{NodeMPST} to support optional \texttt{Promise}s:

@@

Similar changes apply to other types of states in both
\fancyname{NodeMPST} and \fancyname{ReactMPST}.
For \fancyname{ReactMPST}, channel-related functions in the runtime
return the successor state identifier. Hence, aside from
refactoring the functions to delay computation using thunks,
the functions themselves need to
return \texttt{MaybePromise<State>} instead
to propagate the asynchrony
to the execution of the EFSM -- the transition occurs
\textit{after} the developer's handler finishes execution.

>>>>

Asynchronous APIs are commonplace in modern web programming;
typical uses range from
database operations to third-party API calls.
We motivate the need for supporting asynchronous operations
through an authentication protocol in
\cref{subsection:asyncmotivation},
demonstrate how we support asynchronous APIs through extending
the generated APIs (in \cref{subsection:asyncapi}) 
and runtime (in \cref{subsection:asyncruntime}),
and highlight the limitations of our approach in
\cref{subsection:asynclimit}.

@@

The developer's implementation on 
\cref{line:2fachoice1,line:2fachoice2} is likely to query
a database for credential verification.
Database operations are implemented using \textit{asynchronous},
\textit{non-blocking} APIs; it is
recommended that IO-bound operations do not block the main execution,
so a typical JavaScript database API may require a \textit{callback}
to propagate the return value when it is available.
A server endpoint implementation of \cref{lst:2FA}
may resemble the following:

@@

Callbacks are commonplace in dealing with 
asynchronous APIs in JavaScript
\cite{CallbackHell}; we will go over other widely-used
concurrency primitives shortly.
Whatever the case may be, asynchronous developer implementations
using our generated APIs will not type-check as the function
returns immediately (hence, has return type \lstonelinejs{void}),
even though the callback function has the expected return type.
The developer will encounter the following:

@@

The same scenario is also possible for browser-side implementations:
when the user invokes a UI event, the browser-side logic
may first consult a third-party API for additional information
(which is not part of the communication protocol)
before performing a send action; the \textit{Fetch} API \cite{Fetch}
for making API calls is asynchronous.

@@

This means we should support
the concurrency patterns built into TypeScript:

\subparagraph{Callbacks}
Callback-based APIs are higher-order functions -- they take functions
as parameters and invoke them when the asynchronous operation is complete.
For example,
the \texttt{setTimeout(fn, delay)} function available on the browser
waits for \texttt{delay} milliseconds before invoking \texttt{fn()} --
or more precisely, before queuing the execution of \texttt{fn()}.
As a result, the code listing below prints \lstonelinejs{'Second!'}
before \lstonelinejs{'First!'}, even though \texttt{setTimeout}
was called with 0ms delay.

@@

\subparagraph{Promises}
A Promise formalises the ``completion'' callback
and its error handling construct. The developer passes a ``success''
callback and ``error'' callback, conventionally named
``then'' and ``catch'' respectively, to construct a Promise.
This is analogue to the \texttt{Future} interface in Java \cite{JavaFuture}.
The Promise will invoke the correct handler on completion.
The benefits of Promises in our context is that TypeScript allows
us to specify the type of payload that the Promise resolves to.

@@

To support asynchronous implementations in the
generated APIs, we use a type signature that allows
the developer to use \texttt{Promise}s.
We construct a generic union type, \texttt{MaybePromise<T>},
to specify that the type parameter might be wrapped in a Promise,
and leverage type inference in conditional types to
define a generic type, \texttt{FromPromise<T>}, to extract
the wrapped type.

@@

Now we change the type signatures of our generated APIs
to use \texttt{MaybePromise<T>}.
For example, if the handler of a receive state generated by
\fancyname{NodeMPST} must return \texttt{Implementation.S42},
it is now permitted to return a \texttt{Promise} that resolves
to a value of type \texttt{Implementation.S42}.
The same logic applies to other types of EFSM states
in both \fancyname{NodeMPST} and \fancyname{ReactMPST}.

Hence, the asynchronous \trole{Svr} implementation
for the 2FA protocol shown in \cref{lst:new2fa}
now type-checks. Note that, because 
\lstonelinejs{async} functions return \texttt{Promise}s,
we also support this newer piece of syntactic sugar.
Moreover, our generated APIs work with both synchronous
and asynchronous developer implementations transparently,
as is the convention one would expect from modern JavaScript APIs.

@@

Generally, minimal changes are required for the runtime.
The key observation is that the concrete types for the any
handlers are the same -- the payload may be wrapped in a Promise
construct.
To handle potential Promises, we wrap existing logic
as a delayed computation, colloquially known as a \textit{thunk}
\cite{Thunk}.
We show an example in \cref{fig:thunk} of how we
adapt the \texttt{performSend()} function generated by
\fancyname{NodeMPST} to support optional \texttt{Promise}s:

@@

Similar changes apply to other types of states in both
\fancyname{NodeMPST} and \fancyname{ReactMPST}.
For \fancyname{ReactMPST}, channel-related functions in the runtime
return the successor state identifier. Hence, aside from
wrapping the functions to delay computation using thunks,
the functions themselves need to
return \texttt{MaybePromise<State>} instead
to propagate the asynchrony
to the execution of the EFSM -- the transition occurs
\textit{after} the developer's handler finishes execution.

### 2020-06-17 report/sections/server/extensions/error.tex
Report: feedback

We motivate the need for handling a variety exceptions in
modern web programming in \cref{subsection:errormotivation},
demonstrate how we implement a structured error handling
framework through extending the generated APIs and 
runtime in \cref{subsection:errorapi,subsection:errorruntime}
respectively, and explain the limitations of our framework
in \cref{subsection:errorlimit}.

@@

const onClose = ({ target: socket }) => {
	// Wait for the role again
	waiting.add(socketToRole.get(socket));
}

const onSubscribe = (event) => {
	...
	// Update role-WebSocket mapping.
	roleToSocket[role] = socket;
	socketToRole.set(socket, role);
	waiting.delete(role);
	...
}
\end{lstlisting}
\captionof{lstlisting}{Modified Connection Management in
Server Endpoint with Cancellation}
\label{lst:cancelconnectsvr}
\end{figure}

@@

\begin{lstlisting}[language=javascript,tabsize=2]
private processCancellation(role: Roles.All, reason?: any) {
	this.setState({
		elem: this.props.cancellation(role, reason);
	});
}
\end{lstlisting}

@@

\subsection{Caveats with Asynchronous Operations}
Implementing session cancellation whilst
preserving support for asynchronous operations introduces
subtle caveats.

>>>>

We motivate the need for handling a variety exceptions in
modern web programming in \cref{subsection:errormotivation} and
demonstrate how we implement a structured error handling
framework through extending the generated APIs 
(in \cref{subsection:errorapi}) and 
runtime (in \cref{subsection:errorruntime}).
We highlight the caveats of our error handling framework
when dealing with asynchronous operations 
(in \cref{subsection:errorasync}),
and conclude by explaining the limitations of our framework
in \cref{subsection:errorlimit}.

@@

// Invoked when socket is closed
const onClose = ({ target: socket }) => {
	// Wait for the role again
	waiting.add(socketToRole.get(socket));
}

// Invoked when server receives connection request.
const onSubscribe = (event) => {
	...
	// Update role-WebSocket mapping.
	roleToSocket[role] = socket;
	socketToRole.set(socket, role);
	waiting.delete(role);
	...
}
\end{lstlisting}
\captionof{lstlisting}{Modified Connection Management in
Server Endpoint with Cancellation}
\label{lst:cancelconnectsvr}
\end{figure}

@@

\begin{lstlisting}[language=javascript,tabsize=2]
private processCancellation(role: Roles.All, reason?: any) {
	this.setState({
		// Calls the user-specified cancellation handler,
		// which returns a React node to render on DOM
		elem: this.props.cancellation(role, reason);
	});
}
\end{lstlisting}

@@

\subsection{Caveats with Asynchronous Operations}
\label{subsection:errorasync}

Implementing session cancellation whilst
preserving support for asynchronous operations introduces
subtle caveats.
